


\input{header.tex}

%%%%%%%%%%%%%%%%%% TITLE %%%%%%%    %%%%%%%%

\title{\centering The Political Legacy of Displacement: \\ Evidence from the Spanish Exile\footnote{\footnotesize We wish to express our sincere gratitude to Hillel Rapoport, Giovanni Peri, Victor Gay, Andreas Steinmayr, Tanya Surovtseva, 
Mario Carillo, Cédric Chambru, Ángel Herrerín, Paul Maneuvrier-Hervieu, Josepa Miquel-Florensa  and Ricardo Turati, as well as all participants of the 8th Economics and Politics Workshop, the Congrès de l'Association Française d'Histoire Economique. }}
\author{
Cristina Aranzana\footnote{University of Barcelona. Email: \href{mailto:c.aranzana@ub.edu}{c.aranzana@ub.edu}} \and
Luc Paluskiewicz\footnote{Paris School of Economics, Collège de France. Email: \href{mailto:luc.paluskiewicz@psemail.eu}{luc.paluskiewicz@psemail.eu}}
}

\date{ This version: \today \\ \href{https://www.dropbox.com/scl/fi/olo8q2405bpl3xrcra1wg/Draft_Aranzana&Paluskiewicz.pdf?rlkey=61dwvuv15u9m3tsyofifhk8m0&st=ekrs2y9v&dl=0}{[Click here for the latest update]}. }




%%%%%%%%%%%%%%%%%% TEXT %%%%%%%%%%%%%%%

\begin{document}

\maketitle  

\thispagestyle{empty}

\begin{abstract}

\noindent 

This paper studies the long-run political consequences of forced displacement when refugees are politically selected. In 1939, the collapse of the Spanish Republic triggered the sudden arrival of nearly 500,000 left-leaning refugees in France. French authorities, unprepared for the scale of the inflow, interned the displaced population in camps whose locations were determined primarily by logistical constraints. We exploit this quasi-random location of the camps to identify the causal impact of refugee exposure on local political behaviour. Municipalities hosting camps experienced a persistent decline in support for the socialist party, consistent with political backlash. Meanwhile, exposure increased support for the Communist Party and increased the number of individuals involved in the resistance, in line with the diffusion of political ideas. Taken together, the results show that the arrival of politically motivated refugees can generate political polarisation. 

\end{abstract}

\newpage

\setcounter{page}{1}
\section{Introduction}\label{sec:introduction}
\input{sections/introduction.tex}

\section{Historical Context}\label{sec:context}
\input{sections/context.tex}

\section{Data}\label{sec:data}
\input{sections/data.tex}

\section{Descriptive statistics}\label{sec:descriptives}
\input{sections/descriptives.tex}

\section{Empirical Strategy}\label{sec:empirical_strategy}
\input{sections/empirical_strategy.tex}

\section{Preliminary Results}\label{sec:results}
\input{sections/results.tex}

%\clearpage
%\section{Mechanisms}\label{sec:mechanisms}
%\input{sections/mechanisms}

\section{Robustness}\label{sec:robustness}
\input{sections/robustness}

\newpage
\section{Conclusion}\label{sec:conclusion}
\input{sections/conclusion.tex}






%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\begin{singlespace}
%\bibliographystyle{plainnat}
\bibliographystyle{chicago}
\bibliographystyle{aer}
\bibliography{bibliography}
\end{singlespace}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\appendix


\section*{Appendix}\label{sec:appendix}
\input{sections/annex}

\end{document}

